### 2013-04-17 article.tex
Make features clear at start of results
Signed-off-by: Michael Barton <mail@michaelbarton.me.uk>

% Template for PLoS
% Version 1.0 January 2009
%
\documentclass[10pt]{article}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{article}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Genomer - a Swiss army knife for genome scaffolding.}
}
% Insert Author names, affiliations and corresponding author email.
\\
Michael D. Barton$^{1\ast}$, 
Hazel A. Barton$^{1}$
\\
\bf{1} Biology Department, The University of Akron, Akron, OH, 44325-3908, USA
\\
$\ast$ E-mail: mail@michaelbarton.me.uk
\end{flushleft}

\section*{Abstract}

The increasing accessibility and reduced costs of sequencing has made genome
analysis accessible to more and more researchers. In contrast there remains a
steep learning curve in the subsequent computational steps to required to
process raw sequence reads into a database-deposited genome sequence. Here we
describe ``Genomer,'' a tool that simplifies the manual tasks of finishing and
uploading a genome sequence to a database. Genomer can format a genome scaffold
into the common files required for submission to GenBank. Genomer simplifies
editing a genome scaffold by allowing a human-readable YAML format file to be
edited instead of large sequence files. Genomer is written as a command line
tool and is an effort to make the manual, and increasingly common, process of
genome scaffolding more robust and reproducible. Extensive documentation and
video tutorials are available at http://next.gs.

\section*{Introduction}

The decreasing costs and increasing diversity of high-throughput sequencing
methods is making genome analysis a common method to tackle unresolved
questions in microbiology \cite{loman2012b}. Sequencing can produce
$10e^{6}$-$10e^{10}$ short nucleotide reads, which must be assembled into
larger contigs then scaffolded into larger megabase fragments or complete
chromosomes. These larger DNA sequences are then annotated with genomic
features such as protein-coding genes. Each of these required steps can however
produce imperfect results and remains the subject of active research
\cite{earl2011,quail2012,beckloff2012}. Therefore manual curation of a genome
sequence and corresponding annotations may still improve upon the results of
automated methods.

Manually editing a genome sequence or annotation is a non-trivial problem
requiring significant effort on a researcher's part. The large size of FASTA
sequence files or general feature format (GFF) annotation files makes changing
these in a text editor difficult. Tasks such as organising project files,
adding additional sequences to an assembly, updating incorrect annotations or
preparing files for database deposition all divert effort from the main goal of
analysing a genome. As a result genome projects have a steep learning curve for
researchers embarking on genome sequencing for the first time, while still
requiring effort for experienced researchers.

We anticipate these problems have resulted in individual groups independently
producing in-house programming scripts to automate repetitive tasks. An example
of such tasks include finding protein names that do not match a convention of
three lower case characters followed by one upper case character (abcD). Such
tasks are manually laborious but simple to automate computationally. Individual
research groups producing their own scripts for such tasks however leads to a
repetition of effort. In contrast open-source projects can lead to a pooling of
community effort resulting in a higher standard of software and preventing
reinvention of the same code \cite{ince2012}.

During the process of sequencing several \emph{Pseudomonas fluorescens} strains
we developed a tool, called ``Genomer'', to ease and simplify many of the
common steps in our genome project. Several graphical user interface tools are
available to work with genome projects \cite{tanaka2006, wilkinson2002,
lopez2011, carver2012, gordon2003}. Genomer instead provides a simple, command
line interface for genome finishing allowing scriptable automation and
reproducible results. We have open-sourced this tool under the MIT license and
here we describe its application for simplifying steps in a microbial genome
project.

\section*{Implementation}

Genomer is written in the Ruby programming language \cite{ruby-lang,goto2010}
and tested against the 1.8.7 and 1.9.3 versions. Genomer is built upon the
Rubygems and Bundler Ruby package management libraries. Rubygems is a software
management system for Ruby which allows automatic downloading and installation
of third-party software libraries from the rubygems.org website. Bundler is
software for determining which specific versions of the available third
libraries should be used.

Genomer uses these two libraries to create a plugin system where third-party
software can be written and included. Plugin creators need only to prefix the
name of their plugin with \verb+genomer-plugin-+ and upload this to
rubygems.org or make the plugin available as a public git repository. A plugin
is specified within a Genomer project in a file named `Gemfile.' The
\verb+bundle update+ command will then automatically download and install the
required plugins.

Genomer is implemented as a command line tool and tested on both Mac OS X and
Linux systems. The latest version of genomer can installed from the Rubygems
package management system with the command \verb+gem install genomer+. Genomer
is based on our previously described ``scaffolder'' software and is invoked on
the command line with the command \verb+genomer+. Genomer provides extensive
documentation via UNIX manual pages and on the genomer website at
http://next.gs.

\section*{Results and Discussion}

Genomer was developed during our \emph{P. fluorescens} genome projects to
automate and simplify the manual steps required when finishing a microbial
genome. Genomer provides the following functionality to facilitate this:

\begin{description}

  \item[Simple editting of a draft genome sequence.]{Genomer is built on the
  existing Scaffolder \cite{barton2012} file format for assembling a draft
  genome sequence. This YAML \cite{yaml} format simplifies the process of
  re-organising and trimming contigs in a draft genome without having to edit
  large nucleotide sequences manually.}

  \item[Simple mapping of annotations to the generated draft sequence.]{Genomer
  maps the coordinates of annotations on individual to repsective position in
  the draft genome sequence. This allows the scaffold file to be continuously
  improved even when already annotated. For instance, additional PCR sequences
  can be used to close gaps and improve the genome sequence, without requiring
  reannotation or manually updating all annotation coordinates.}

  \item[Generation of files for submission to GenBank.]{Submitting a genome
  sequence to GenBank database requires several specific files to be prepared.
  These may include a FASTA file of the draft sequence, a table of annotations,
  a FASTA file containing the individual contig sequences, and `a golden path'
  (AGP) \cite{agp-spec} file describing the placement of contigs. Genomer
  automates the generation of all these files.}

  \item[Integration with the command line.]{Genomer is built as a command line
  tool built around plain text files following the Linux philosophy. This
  allows for integration with tools such as GNU Make or git, and also to allow
  reproducible scripting of the process of finishing a genome.}

\end{description}

This software was written for the command-line to use in shells such as bash or
zsh. Command line tools are simple to automate using scripting, allowing
reproducible genome finishing and sharing genome projects between researchers.
For instance, during our microbial genome project we automated Genomer using
GNU Make build files, allowing the finishing steps to be repeated automatically
using the \verb+make+ command. Additionally we stored each version of the GNU
Make build file with git \cite{git-scm} to track changes to the project and to
allow errors to be reverted from earlier versions. Example build files from our
genome projects can be found on GitHub for a simple plasmid
\cite{plasmid-github} and a more complex genome \cite{genome-github} sequence.

We divided this tool into modular parts each called a `plugin.' Genomer
provides no functionality itself, but rather serves to make the functionality
of plugins available for use at the command line. For example to validate the
correctness of genome annotations, the genomer-plugin-validate plugin should be
installed which then allows the required functionality to be through the
\verb+genomer+ command. Presently, we have developed and released three
plugins: viewing a genome sequence in common bioinformatics file formats,
validating the correctness of genome annotations for common errors, and provide
summaries of a genome sequence. Each of these plugins are available on GitHub
and their use described is in the documentation at http://next.gs.

The Genomer plugin system was built using the existing Ruby package management
system: RubyGems and Bundler. This allows plugins to be automatically installed
without requiring manual downloading and compilation by the user. We anticipate
this will eliminate a common problem where software may require technical
expertise to first compile and install before being used. A plugin system also
allows each plugin to be developed on different release schedules without
requiring Genomer itself to be updated, thereby providing a stable interface.
Each Genomer plugin can be locked to specific versions to prevent changes in
functionality between different plugin versions. This addresses situations
whereby backwards incompatible software changes can break an existing project's
workflow.

We have released Genomer as open source on GitHub \cite{genomer-github} and
have prepared documentation and video tutorials at http://next.gs. As we have
used this tool extensively to simplify our own genome projects we believe
genomer may also be useful to others in the field. We will illustrate the
possible usefulness of genomer with two use cases: updating the annotations of
a genome scaffold during the finishing process, and preparing genome files for
submission to GenBank.

\subsection*{First Use Case: Continuous Genome Finishing}

Genome annotation is required to identify the features responsible for
phenotype. Online tools such as the Integrated Microbial Genomes system
\cite{markowitz2006} provide an online service whereby a nucleotide sequence
may be submitted and a subsequent file of predicted genomic features returned.
These annotations are required when submitting a genome sequence to a database.
Once annotated modifying the sequence becomes problematic as this requires
corresponding changes in all downstream annotated coordinates. This therefore
presents a problem whereby either no more updates can be made to the sequence
or the delta difference in annotation coordinates must be propagated to all
downstream locations. The former option is often chosen as being simpler and
therefore the genome sequence must be considered `frozen' once annotated and no
more changes can be made.

During our own \emph{P. fluorescens} genome sequencing projects, we aimed to
continuously improve the quality of the genome sequence by closing gaps using
PCR and Sanger sequencing while at the same time analysing the observed genomic
features. We therefore incorporated functionality into Genomer to allow
annotation coordinates to be automatically updated following such changes to
the sequence. Genomer allows this by organising a genome project as follows:

\begin{enumerate}

\item{A genomer project is organised around a set of already assembled and
annotated contigs. For example we assemble our contigs using the A5 pipeline
\cite{tritt2012} and and then subsequently annotate them using the Integrated
Microbial Genomes resource \cite{markowitz2006}. We then use these contigs as a
starting point to build a draft genome sequence in genomer.}

\item{A genome scaffold is then created using the scaffolder file format
\cite{barton2012}. This specifies the order and orientation of contigs in the
scaffold allowing any unresolved gap regions to also be specified. Paired-end
sequencing, comparison to reference genomes, or PCR amplification can be used
to estimate the positions of observed contigs in the true genome sequence.}

\item{The genomer-plugin-view plugin is then added to the Gemfile in the root
of the project directory. The command \verb+bundle update+, run in the project
root, will then download and install this plugin making it available for use in
the Genomer project.}

\item{Different views can then be generated from the scaffold using the
\verb+genomer view+ command. For instance, a GFF3 file of the scaffolded
annotations may be generated. Should the originating contig annotations or the
genome scaffold be changed then rerunning the \verb+genomer view+ command will
correspondingly update the annotation coordinates for the scaffold.}

\end{enumerate}

\subsection*{Second Use Case: Preparing files for submission to GenBank}

The genomer-view-plugin can be used to generate for submission to GenBank:
assembled FASTA, contig FASTA, annotation tables, and scaffold layout in AGP.
This includes FASTA sequences for both the scaffolded and individual contigs, a
GenBank table file of contig annotations, and an AGP file for the order of
contigs in the scaffold. Additional options may also be used to add a prefix to
gene annotation IDs, or begin locus ID numbering at the origin of the sequence,
thereby simplifying the process of creating the required files for submission.

Genomer does not produce the .asn file which must be submitted to GenBank,
instead the generated files are used as the input for the tbl2asn or sequin
tools. We emphasise that the purpose of Genomer is not to reinvent the
functionality already provided by current NCBI resources, but instead to
simplify the process of producing the required inputs. Our example \emph{P.
fluorescens} GNU Make file \cite{genome-github} illustrates how Genomer may be
combined with tbl2asn to produce the required .asn file for submission.

\subsection*{Summary}

Genomer is an open-source tool to simplify and automate repetitive and
time-consuming tasks required when finishing a microbial genome project. This
software is available on GitHub \cite{genomer-github} with documentation and
video tutorials on the http://next.gs website. This tool has been useful in our
own genome projects and we believe it will also be useful to other researchers
especially smaller research groups entering the field for the first time.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgements}

This work was supported by the National Institute for Health: IDeA Network of
Biomedical Research Excellence (KY-INBRE) grant (NIH 2P20 RR016481-09) and the
NIH R15 AREA Program grant (R15GM079775).

%\section*{References}
% The bibtex filename
\bibliography{article}

\end{document}